Structure constants and short-time simulations of open collective quantum spin systems

### 2019-03-28 soc_squeezing/collective_paper/collective_spin.tex
added some text/notes on two-time correlators commented out the last appendix on more general fixed-weight operator products

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
other spins by $s_n^{(j)}$.  We then define the collective spin
operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
collective spin operators in the form
\begin{align}
  \O = \sum_{\v m} \O_{\v m} \S_{\v m}.
\end{align}
The time evolution of a Heisenberg operator $\O\p{t}$ under a
Hamiltonian $H$ is then given by
\begin{align}
  \f{d}{dt} \O\p{t}
  = T \O\p{t}
  = \sum_{\v m, \v n} \S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:EOM}
\end{align}
for a time derivative (super-)operator $T$ with matrix elements
$T_{\v m\v n}$, determined by
\begin{align}
  T \O \equiv i \sp{H, \O}_- + \sum_\J \gamma_\J \D\p{\J; \O},
  \label{eq:time_deriv}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
with a corresponding decoherence rate $\gamma_\J$; and
\begin{align}
  \D\p{\J;\O}
  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
\end{align}
is the Heisenberg dissipator, or Lindblad superoperator, for all jump
operators $J\in\J$.  Decoherence via uncorrelated decay of individual
spins at a rate $\gamma_-$, for example, would be described by the set
of jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$ with
$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:time_deriv}
can be determined from the structure constants $f_{\v\ell\v m\v n}$ of
the general product
$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ worked
out in Appendices \ref{sec:identities}--\ref{sec:prod_general}, while
the effects of decoherence from jump operators (i.e.~elements of $\J$)
of the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
$\Gamma=\sum_n\Gamma_nS_n$ are worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
consider these calculations to be the main technical contributions of
our work, with potential applications beyond the short-time simulation
method presented here.  These ingredients are sufficient to compute
matrix elements $T_{\v m\v n}$ of the time derivative operator $T$ in
Eq.~\eqref{eq:time_deriv} in most cases of practical interest.

The time derivative operator $T$ will generally couple spin operators
$\S_{\v n}$ (i.e.~in $\O$) to spin operators $\S_{\v m}$ with higher
``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
weight signifies the growth of many-body correlations.  In practice,
keeping track of this growth will eventually require more
computational resources than are available, meaning we must somehow
truncate our equations of motion.  The simplest truncation strategy
would be to take
\begin{align}
  \f{d}{dt} \O
  \to \sum_{w\p{\v m}<W} \S_{\v m}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and cutoff
$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
system of differential equations defined by Eq.~\eqref{eq:EOM}, and
allows us to solve it using standard numerical methods.  Initial
conditions for this system of differential equations, i.e.~expectation
values of collective spin operators with respect to states that are
simple to prepare experimentally, are provided in Appendix
\ref{sec:initial_conditions}.

@@

We can expand Heisenberg operators $\O\p{t}$ in a Taylor series about
the time $t=0$ to write
\begin{align}
  \O\p{t}
  = e^{t T} \O\p{0}
  % = \sum_{k\ge0} \f{t^k}{k!} T^k \O\p{0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
  \label{eq:time_series}
\end{align}
where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
derivative are
\begin{align}
  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
  T^{k>1}_{\v m\v n}
  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
  T_{\v p_2\v p_1} T_{\v p_1\v n},
\end{align}
with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
sufficiently short times, we can truncate the series in
Eq.~\eqref{eq:time_series} by taking
\begin{align}
  \O\p{t}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
  \label{eq:order_truncation}
\end{align}
Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
which operators are relevant for computing the expectation value
$\bk{\O\p{t}}$ to a fixed order $M$.  Note that the relation
$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$, which
by Hermitian conjugation of Eq.~\eqref{eq:EOM} also implies that
$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
number of initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and
the number of matrix elements $T_{\v m\v n}$ that we may need to
explicitly compute roughly in half.

@@

\section{Two-time correlation functions}
\label{sec:TTC}

As a final example, we consider the calculation of two-time
correlation functions in collective spin systems.

@@

\section{Changing operator bases, and a general product for algebras
  of permutationally-symmetric operators}
\label{sec:new_basis}

For a collective spin system composed of many small spins, the
operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
$\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
but also carry lots of ``baggage'' in the form of $k$-body operators
with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
our basis for collective spin operators from $\S_{\v m}$ to the
purely-$\abs{\v m}$-body operators
\begin{align}
  \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
  &&
  P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
  \label{eq:S_P}
\end{align}
where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
ordered list of indices for the spins addressed by $P_{\v j}$; and the
sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
which all indices in $\v j$ are distinct and
$\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
spin addressed by an $s_\mu$ operator in $P_{\v j}$.

In order to make use of the operators $\tilde S_{\v m}$, we will need
to compute products of the form $\tilde S_{\v m} \tilde S_{\v n}$.  We
therefore expand such a product into terms that have exactly $s$ spins
addressed by both operators:
  \tilde\S_{\v m} \tilde\S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
  \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
  \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
      \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X.
  \label{eq:SS_PP}
Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
$P_{\v j}$ address the same spin as an $s_\nu$ operator in $P_{\v k}$,
we have
  \tilde\S_{\v m} \tilde\S_{\v n}
  = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
  \sum_{\p{\v J,\v K;\v m,\v n,\v r}} P_{\v J} Q_{\v K},
  \label{eq:SS_PQ}
where the sum over $\p{\v r;\v m,\v n,s}$ denotes a sum over all
values of $\v r$ with the restrictions
  r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
  \sum_\nu r_{\mu\nu} \le m_\mu,
  \sum_\mu r_{\mu\nu} \le n_\nu,
  \sum_{\mu,\nu} r_{\mu\nu} = s;
  \label{eq:rest_r}
the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
operators in $P_{\v j}$ and $P_{\v k}$ according to $\v r$, or
  f_{\v m\v n\v r}
  &\equiv \sp{\prod_{\mu,\nu}
    { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
    { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
  &= \sp{\prod_\mu \f{m_\mu!}{\p{m_\mu-\sum_\rho r_{\mu\rho}}!}
    \f{n_\mu!}{\p{n_\mu-\sum_\rho r_{\rho\mu}}!}}
  \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1};
$\v J$ and $\v K$ are respectively ordered lists of indices for spins
addressed by one and two single-spin operators in a term of the
product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
$\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
$\v J,\v K$ with the restrictions
  \abs{\set{J_a^\mu}}
  = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
  \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
  \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \varnothing,
with $\varnothing$ denoting the empty set; and finally, similarly to
$P_{\v j}$ in \eqref{eq:S_P} we define
  Q_{\v K} \equiv \prod_{\mu,\nu,a}
  s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}.
As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
indices in $\v K$, we can safely neglect keeping track of these
indices and simply write
  Q_{\v K}
  = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
  s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}
  \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}} s_\mu s_\nu
  = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
  \sum_\rho \eta_{\mu\nu\rho} s_\rho,
  \label{eq:Q_K_eta}
where $\simeq$ denotes equality up to a re-indexing of spins and
additional tensor factors of the single-spin identity operator
(i.e.~$\1$); we have introduced explicit dependence on
$r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
$\eta_{\mu\nu\rho}$ is a structure constant defined by
$s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
$\mu,\nu$ in most of the work above, the sum over $\rho$ here includes
an index for the identity operator $s_0\equiv\1$.  Distributing the
product of sums into a sum of products gives us
  Q_{\v K}
  \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
  \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
  = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}\eta_{\mu\nu\rho^{\mu\nu}_a}}
  \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
  \label{eq:Q_K_rho}
where implicitly $\abs{\set{\rho^{\mu\nu}_a}}=r_{\mu\nu}$, as
$\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta} and
$\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
$\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
index list
$\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
that are equal to $\kappa$, we observe that two terms in
\eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
are equal up to a permutation of indices if
$\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy (under
permutation) of terms in \eqref{eq:Q_K_rho}, i.e.~the number of
$\v\rho$ that are consistent with $\tilde{\v\rho}$, is
  g_{\tilde{\v\rho}}
  \equiv \prod_{\mu,\nu,\kappa}
  { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
    \choose \tilde\rho^{\mu\nu}_\kappa }
  = \sp{\prod_{\mu,\nu} r_{\mu\nu}!}
  \sp{\prod_{\mu,\nu,\kappa} \tilde\rho^{\mu\nu}_\kappa!}^{-1}.
Defining additionally
  \v\eta^{\tilde{\v\rho}}
  \equiv \prod_{\mu,\nu,\kappa}
  \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
  \label{eq:eta_rho}
we can collect like factors and group together equivalent terms in
\eqref{eq:Q_K_rho} to write
  Q_{\v K}
  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
  \bigotimes_{\kappa}
  s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
  \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
  \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
  \otimes \bigotimes_{\kappa\ne0}
  s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
  \label{eq:Q_K_rho_simp}
where again implicitly
$\sum_\kappa\tilde\rho^{\mu\nu}_\kappa=r_{\mu\nu}$ for consistency
with \eqref{eq:Q_K_eta} and \eqref{eq:Q_K_rho}, and we have explicitly
factored out the identity operators in each term of $Q_{\v K}$.

>>>>

In this section we provide the basic theory for our method to compute
expectation values of collective spin operators, deferring lengthy
derivations to the appendices.  We consider systems composed of $N$
distinct spin-1/2 particles.  Defining individual spin-1/2 operators
$s_{n=\x,\y,\z}\equiv\sigma_n/2$ and
$s_\pm\equiv s_\x\pm is_\y=\sigma_\pm$ with Pauli operators
$\sigma_n$, we denote an operator that acts with $s_n$ on the spin
indexed by $j$ and trivially (i.e.~with the identity $\1$) on all
other spins by $s_n^{(j)}$.  We then define the collective spin
operators $S_n\equiv\sum_{j=1}^Ns_n^{(j)}$ for
$n\in\set{\x,\y,\z,+,-}$.  Identifying the set $\set{\S_{\v m}}$ as a
basis for all collective spin operators, with
$\v m\equiv\p{m_+,m_\z,m_-}\in\N_0^3$ and
$\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$, we can expand all
collective spin operators in the form
\begin{align}
  \O = \sum_{\v m} \O_{\v m} \S_{\v m}.
\end{align}
The time evolution of a Heisenberg operator $\O\p{t}$ under a
Hamiltonian $H$ is then given by
\begin{align}
  \f{d}{dt} \O\p{t}
  = T \O\p{t}
  = \sum_{\v m, \v n} \S_{\v m} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:EOM}
\end{align}
for a time derivative (super-)operator $T$ with matrix elements
$T_{\v m\v n}$, determined by
\begin{align}
  T \O \equiv i \sp{H, \O}_- + \sum_\J \gamma_\J \D\p{\J; \O},
  \label{eq:time_deriv}
\end{align}
where $\sp{X,Y}_\pm\equiv XY\pm YX$; $\J$ is a set of jump operators
with a corresponding decoherence rate $\gamma_\J$; and
\begin{align}
  \D\p{\J;\O}
  \equiv \sum_{J\in\J} \p{J^\dag \O J - \f12\sp{J^\dag J,\O}_+}
\end{align}
is the Heisenberg dissipator, or Lindblad superoperator, for all jump
operators $J\in\J$.  Decoherence via uncorrelated decay of individual
spins at a rate $\gamma_-$, for example, would be described by the set
of jump operators $\J_-=\set{s_-^{(j)}:j=1,2,\cdots,N}$ with
$\gamma_{\J_-}=\gamma_-$.  The commutator in Eq.~\eqref{eq:time_deriv}
can be computed by expanding the product
$\S_{\v\ell}\S_{\v m}=\sum_{\v n}f_{\v\ell\v m\v n}\S_{\v n}$ with
structure constants $f_{\v\ell\v m\v n}$ that are worked out in
Appendices \ref{sec:identities}--\ref{sec:prod_general}, and the
effects of decoherence from jump operators (i.e.~elements of $\J$) of
the form $\gamma^{(j)}=\sum_n\gamma_ns_n^{(j)}$ and
$\Gamma=\sum_n\Gamma_nS_n$ are worked out in Appendices
\ref{sec:sandwich_single}--\ref{sec:decoherence_collective}.  We
consider these calculations to be the main technical contributions of
our work, with potential applications beyond the short-time simulation
method presented here.  These ingredients are sufficient to compute
matrix elements $T_{\v m\v n}$ of the time derivative operator $T$ in
Eq.~\eqref{eq:time_deriv} in most cases of practical interest.

The time derivative operator $T$ will generally couple spin operators
$\S_{\v n}$ (i.e.~in $\O$) to spin operators $\S_{\v m}$ with higher
``weight'', i.e.~with $\abs{\v m}>\abs{\v n}$, where
$\abs{\v\ell}\equiv\sum_\alpha\ell_\alpha$.  The growth of operator
weight signifies the growth of many-body correlations.  In practice,
keeping track of this growth will eventually require more
computational resources than are available, meaning we must somehow
truncate our equations of motion.  The simplest truncation strategy
would be to take
\begin{align}
  \f{d}{dt} \O\p{t}
  \to \sum_{w\p{\v m}<W} \S_{\v m}
  \sum_{\v n} T_{\v m\v n} \O_{\v n}\p{t}
  \label{eq:weight_truncation}
\end{align}
for some weight measure $w$, e.g.~$w\p{\v m}=\abs{\v m}$, and cutoff
$W$.  The truncation in Eq.~\eqref{eq:weight_truncation} closes the
system of differential equations defined by Eq.~\eqref{eq:EOM}, and
allows us to solve it using standard numerical methods.  Initial
conditions for this system of differential equations, i.e.~expectation
values of collective spin operators with respect to states that are
simple to prepare experimentally, are provided in Appendix
\ref{sec:initial_conditions}.

@@

We can expand Heisenberg operators $\O\p{t}$ in a Taylor series about
the time $t=0$ to write
\begin{align}
  \O\p{t}
  = e^{t T} \O\p{0}
  = \sum_{k\ge0} \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0},
  \label{eq:time_series}
\end{align}
where the matrix elements $T^k_{\v m\v n}$ of the $k$-th time
derivative operator $T^k$ are
\begin{align}
  T^0_{\v m\v n} &\equiv \delta_{\v m\v n}, \\
  T^1_{\v m\v n} &\equiv T_{\v m\v n}, \\
  T^{k>1}_{\v m\v n}
  &\equiv \sum_{\v p_1,\v p_2,\cdots,\v p_{k-1}}
  T_{\v m\v p_{k-1}} \cdots T_{\v p_3\v p_2}
  T_{\v p_2\v p_1} T_{\v p_1\v n},
\end{align}
with $\delta_{\v m\v n}=1$ if $\v m=\v n$ and 0 otherwise.  For
sufficiently short times, we can truncate the series in
Eq.~\eqref{eq:time_series} by taking
\begin{align}
  \O\p{t}
  \to \sum_{k=0}^M \f{t^k}{k!}
  \sum_{\v m, \v n} \S_{\v m} T^k_{\v m\v n} \O_{\v n}\p{0}.
  \label{eq:order_truncation}
\end{align}
Unlike the truncation in Eq.~\eqref{eq:weight_truncation}, the nonzero
matrix elements $T^k_{\v m\v n}$ for $k=0,1,\cdots,M$ now tell us
which operators are relevant for computing the expectation value
$\bk{\O\p{t}}$ to a fixed order $M$.  Note that the relation
$\S_{\v m}^\dag=\S_{\v m^*}$ for $\v m^*\equiv\p{m_-,m_\z,m_+}$, which
by Hermitian conjugation of Eq.~\eqref{eq:EOM} also implies that
$T_{\v m^*\v n^*}=T_{\v m\v n}^*$, can be exploited to cut both the
number of initial-time expectation values $\bk{\S_{\v m}}_{t=0}$ and
the number of matrix elements $T_{\v m\v n}$ that we may need to
explicitly compute roughly in half.

@@

\section{Two-time correlation functions and non-equilibrium linear
  response}
\label{sec:TTC}

As a final example, we consider the calculation of two-time
correlation functions in a collective spin system.  These correlation
functions allow us to extract the short-time linear response of
Heisenberg operators to small perturbations of a collective spin
system's Hamiltonian.  Consider a target Hamiltonian $H$ perturbed by
an operator $V$, such that the net Hamiltonian is $\widetilde H=H+V$.
We denote the generator of Heisenberg time evolution under the
perturbed (unperturbed) Hamiltonian by $\widetilde T$ ($T$).  These
time evolution operators are related by
\begin{align}
  \widetilde T = T + \hat V
\end{align}
where $\hat V$ is a superoperator whose action on operators $\O$ is
defined by
\begin{align}
  \hat V \O \equiv i\sp{V,\O}_-.
\end{align}
By Taylor expanding through quadratic order in the time $t$, we can
find that for sufficiently small times
\begin{align}
  e^{t\hat V} e^{t T}
  \approx e^{t\widetilde T} - \f12 t^2 \sp{T,\hat V}_-.
\end{align}
To leading order in $\hat V$ and quadratic order in $t$, we therefore
have
\begin{align}
  e^{t\widetilde T} - e^{tT}
  \approx t\hat V e^{tT} + \f12 t^2 \sp{T,\hat V}_-,
\end{align}
where
\begin{align}
  \f12 t^2 \sp{T,\hat V}_- \approx \f12 t \sp{e^{tT}, \hat V}_-,
\end{align}
which implies
\begin{align}
  e^{t\widetilde T} - e^{tT} \approx \f12 t \sp{e^{tT},\hat V}_+.
\end{align}
Defining the perturbed and unperturbed Heisenberg operators
$\widetilde\O\p{t}\equiv e^{t\widetilde T}\O$ and
$\O\p{t}\equiv e^{tT}\O$, we thus find
\begin{align}
  \widetilde\O\p{t} - \O\p{t}
 \approx \f12 i t \p{\sp{V,\O\p{t}}_- + \sp{V,\O}_-\p{t}}.
\end{align}

\newpage

@@

% \section{Changing operator bases, and a general product for algebras
%   of permutationally-symmetric operators}
% \label{sec:new_basis}

% For a collective spin system composed of many small spins, the
% operators $\S_{\v m}\equiv S_+^{m_+} S_\z^{m_\z} S_-^{m_-}$ contain
% $\abs{\v m}$-body operators (for $\abs{\v m}\equiv\sum_\mu m_\mu$),
% but also carry lots of ``baggage'' in the form of $k$-body operators
% with $k<\abs{\v m}$.  To cut down on this overhead, we try to change
% our basis for collective spin operators from $\S_{\v m}$ to the
% purely-$\abs{\v m}$-body operators
% \begin{align}
%   \tilde\S_{\v m} \equiv \sum_{\p{\v j;\v m}} P_{\v j},
%   &&
%   P_{\v j} \equiv \prod_{\mu,a} s_\mu^{\p{j^\mu_a}},
%   \label{eq:S_P}
% \end{align}
% where $\v m \equiv \p{m_+,m_\z,m_-}\in\mathbb{N}_0^3$ with $m_\mu$ the
% number of $s_\mu$ operators in each term of $\S_{\v m}$; $\v j$ is an
% ordered list of indices for the spins addressed by $P_{\v j}$; and the
% sum over $\p{\v j;\v m}$ denotes a sum over all possible $\v j$ for
% which all indices in $\v j$ are distinct and
% $\abs{\set{j^\mu_a}} = m_\mu$, such that $j^\mu_a$ indexes the $a$-th
% spin addressed by an $s_\mu$ operator in $P_{\v j}$.

% In order to make use of the operators $\tilde S_{\v m}$, we will need
% to compute products of the form $\tilde S_{\v m} \tilde S_{\v n}$.  We
% therefore expand such a product into terms that have exactly $s$ spins
% addressed by both operators:
% \begin{align}
%   \tilde\S_{\v m} \tilde\S_{\v n}
%   = \sum_{s\ge0} \sum_{\p{\v j,\v k;\v m,\v n,s}} P_{\v j} P_{\v k},
%   &&
%   \sum_{\p{\v j,\v k;\v m,\v n,s}} X \equiv
%   \sum_{\substack{\p{\v j;\v m},\p{\v k;\v n} \\
%       \abs{\set{j_\alpha^\mu}\cap\set{k_\beta^\nu}}=s}} X.
%   \label{eq:SS_PP}
% \end{align}
% Collecting terms in which $r_{\mu\nu}$ of the $s_\mu$ operators in
% $P_{\v j}$ address the same spin as an $s_\nu$ operator in $P_{\v k}$,
% we have
% \begin{align}
%   \tilde\S_{\v m} \tilde\S_{\v n}
%   = \sum_{s\ge0} \sum_{\p{\v r;\v m,\v n,s}} f_{\v m\v n\v r}
%   \sum_{\p{\v J,\v K;\v m,\v n,\v r}} P_{\v J} Q_{\v K},
%   \label{eq:SS_PQ}
% \end{align}
% where the sum over $\p{\v r;\v m,\v n,s}$ denotes a sum over all
% values of $\v r$ with the restrictions
% \begin{align}
%   r_{\mu\nu} \ge 0 ~\forall~ \mu,\nu,
%   &&
%   \sum_\nu r_{\mu\nu} \le m_\mu,
%   &&
%   \sum_\mu r_{\mu\nu} \le n_\nu,
%   &&
%   \sum_{\mu,\nu} r_{\mu\nu} = s;
%   \label{eq:rest_r}
% \end{align}
% the factor $f_{\v m\v n\v r}$ counts the number of ways to pair
% operators in $P_{\v j}$ and $P_{\v k}$ according to $\v r$, or
% \begin{align}
%   f_{\v m\v n\v r}
%   &\equiv \sp{\prod_{\mu,\nu}
%     { m_\mu - \sum_{\rho<\nu} r_{\mu\rho} \choose r_{\mu\nu} }
%     { n_\mu - \sum_{\rho<\nu} r_{\rho\mu} \choose r_{\nu\mu} }}
%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!} \\
%   &= \sp{\prod_\mu \f{m_\mu!}{\p{m_\mu-\sum_\rho r_{\mu\rho}}!}
%     \f{n_\mu!}{\p{n_\mu-\sum_\rho r_{\rho\mu}}!}}
%   \sp{\prod_{\mu,\nu} r_{\mu\nu}!}^{-1};
% \end{align}
% $\v J$ and $\v K$ are respectively ordered lists of indices for spins
% addressed by one and two single-spin operators in a term of the
% product $P_{\v j} P_{\v k}$ with fixed $\v r$; the sum over
% $\p{\v J,\v K;\v m,\v n,\v r}$ denotes a sum over all values of
% $\v J,\v K$ with the restrictions
% \begin{align}
%   \abs{\set{J_a^\mu}}
%   = m_\mu + n_\mu - \sum_\rho\p{r_{\mu\rho}+r_{\rho\mu}},
%   &&
%   \abs{\set{K^{\mu\nu}_a}} = r_{\mu\nu},
%   &&
%   \set{J_a^\lambda} \cap \set{K_b^{\mu\nu}} = \varnothing,
% \end{align}
% with $\varnothing$ denoting the empty set; and finally, similarly to
% $P_{\v j}$ in \eqref{eq:S_P} we define
% \begin{align}
%   Q_{\v K} \equiv \prod_{\mu,\nu,a}
%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}.
% \end{align}
% As the sum in \eqref{eq:SS_PQ} is invariant under permutation of the
% indices in $\v K$, we can safely neglect keeping track of these
% indices and simply write
% \begin{align}
%   Q_{\v K}
%   = \prod_{\mu,\nu} \prod_{a=1}^{r_{\mu\nu}}
%   s_\mu^{(K^{\mu\nu}_a)} s_\nu^{(K^{\mu\nu}_a)}
%   \simeq \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}} s_\mu s_\nu
%   = \bigotimes_{\mu,\nu} \bigotimes_{a=1}^{r_{\mu\nu}}
%   \sum_\rho \eta_{\mu\nu\rho} s_\rho,
%   \label{eq:Q_K_eta}
% \end{align}
% where $\simeq$ denotes equality up to a re-indexing of spins and
% additional tensor factors of the single-spin identity operator
% (i.e.~$\1$); we have introduced explicit dependence on
% $r_{\mu\nu}=\abs{\set{K^{\mu\nu}_a}}$ for brevity; and
% $\eta_{\mu\nu\rho}$ is a structure constant defined by
% $s_\mu s_\nu=\sum_\rho\eta_{\mu\nu\rho}s_\rho$.  Unlike the sums over
% $\mu,\nu$ in most of the work above, the sum over $\rho$ here includes
% an index for the identity operator $s_0\equiv\1$.  Distributing the
% product of sums into a sum of products gives us
% \begin{align}
%   Q_{\v K}
%   \simeq \sum_{\v\rho} \bigotimes_{\mu,\nu,a}
%   \eta_{\mu\nu\rho^{\mu\nu}_a} s_{\rho^{\mu\nu}_a}
%   = \sum_{\v\rho} \sp{\prod_{\mu,\nu,a}\eta_{\mu\nu\rho^{\mu\nu}_a}}
%   \bigotimes_{\mu,\nu,a} s_{\rho^{\mu\nu}_a},
%   \label{eq:Q_K_rho}
% \end{align}
% where implicitly $\abs{\set{\rho^{\mu\nu}_a}}=r_{\mu\nu}$, as
% $\p{\mu,\nu,a}$ essentially index a factor in \eqref{eq:Q_K_eta} and
% $\rho^{\mu\nu}_a$ indexes one of the terms in that factor.  Letting
% $\tilde\rho^{\mu\nu}_\kappa$ denote the number of elements in the
% index list
% $\p{\rho^{\mu\nu}_1,\rho^{\mu\nu}_2,\cdots,\rho^{\mu\nu}_{r_{\mu\nu}}}$
% that are equal to $\kappa$, we observe that two terms in
% \eqref{eq:Q_K_rho} with, say, $\v\rho=\v\rho_1$ and $\v\rho=\v\rho_2$
% are equal up to a permutation of indices if
% $\tilde{\v\rho}_1=\tilde{\v\rho}_2$.  The degeneracy (under
% permutation) of terms in \eqref{eq:Q_K_rho}, i.e.~the number of
% $\v\rho$ that are consistent with $\tilde{\v\rho}$, is
% \begin{align}
%   g_{\tilde{\v\rho}}
%   \equiv \prod_{\mu,\nu,\kappa}
%   { r_{\mu\nu} - \sum_{\lambda<\kappa} \tilde\rho^{\mu\nu}_\lambda
%     \choose \tilde\rho^{\mu\nu}_\kappa }
%   = \sp{\prod_{\mu,\nu} r_{\mu\nu}!}
%   \sp{\prod_{\mu,\nu,\kappa} \tilde\rho^{\mu\nu}_\kappa!}^{-1}.
% \end{align}
% Defining additionally
% \begin{align}
%   \v\eta^{\tilde{\v\rho}}
%   \equiv \prod_{\mu,\nu,\kappa}
%   \p{\eta_{\mu\nu\kappa}}^{\tilde\rho^{\mu\nu}_\kappa},
%   \label{eq:eta_rho}
% \end{align}
% we can collect like factors and group together equivalent terms in
% \eqref{eq:Q_K_rho} to write
% \begin{align}
%   Q_{\v K}
%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
%   \bigotimes_{\kappa}
%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa}
%   \simeq \sum_{\tilde{\v\rho}} g_{\tilde{\v\rho}} \v\eta^{\tilde{\v\rho}}
%   \1^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_0}
%   \otimes \bigotimes_{\kappa\ne0}
%   s_\kappa^{\otimes\sum_{\mu,\nu}\tilde\rho^{\mu\nu}_\kappa},
%   \label{eq:Q_K_rho_simp}
% \end{align}
% where again implicitly
